---
title: "Lab3Che"
author: "Che Hoon Jeong"
date: "2/12/2022"
output:
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
    code_folding: hide
---

### Question 1

There are many different hieroglyphs in the file, including birds, geometric shapes, tools, stars, rabbits, etc. The rough estimate is about 25. 

### Question 2

We read in the hieroglyph images and store them as a single data frame. We also change the images' shape, so that the dataframe has 4410 rows (images) and 3740 columns (pixels).

```{r, eval=FALSE, warning=FALSE, error=FALSE}

#reading in all the images as single dataframe

setwd("/Users/CheHoon/Desktop/SP2022/DA350/Lab3/Hieroglyphics_Images")
files = list.files()
hieroglyphs <- lapply(files, function(x) readPNG(x))
imagesdf <- data.frame(do.call(rbind, lapply(hieroglyphs, function(x) t(matrix(x)))))

#we export the dataframe as csv, to save computation time
write.csv(imagesdf, "/Users/CheHoon/Desktop/SP2022/DA350/Lab3/images.csv")

imagesdf <- read.csv("/Users/CheHoon/Desktop/SP2022/DA350/Lab3/images.csv")


```


### Question 3

We create principal components, and generate graph to obtain number of components that explain 95% of variance in the data.

```{r, eval=FALSE, warning=FALSE, error=FALSE}
#we compress the data using PCA.
images.pca = prcomp(imagesdf[,], scale.=TRUE)

##Plotting proportion of variance explained
pca_var = images.pca$sdev^2
prop_varex <- pca_var/sum(pca_var) #Proportion of variance
plot(cumsum(prop_varex),type='b', main = "Cumulative Proportion of Variance") #Cumulative proportion of variance
cumsum(prop_varex)


```

We can see that 557 Components explain 95% of the variance in the data.


### Question 4

We run k-means clustering algorithms on the principal component data with varying number of cluster numbers.

```{r, eval=FALSE, warning=FALSE, error=FALSE}

clusterNum <- 200

clusterModels <- sapply(1:clusterNum, function(i) kmeans(images.pca$x[,1:557],i))
wcs <- sapply(1:clusterNum, function(i) sum(clusterModels[,i]$tot.withinss))

plot(1:clusterNum, wcs, type="b", xlab="Number of Clusters",
     ylab="Within Cluster Sum of Squares")



```

We see that 25 is roughly the elbow of the plot.


### Question 5.

We plot a histogram of the within cluster sum of squares.

```{r, eval=FALSE, warning=FALSE, error=FALSE}

#Plot within-cluster distances as a function of cluster size
wss <- sapply(1:1000, function(k) kmeans(images.pca$x[,1:557], 25, nstart=1)$tot.withinss)
hist(wss, breaks = 50)

```

We see that the most frequent count is around 90 which has a fair within sum square. We therefore perform just a little more random start, so that we have a fair chance of obtaining a good solution.

```{r, eval=FALSE, warning=FALSE, error=FALSE}

myCluster <- kmeans(images.pca$x[,1:557], 25, nstart=100)

```



### Question 6.

We put the hieroglyph images into their designated clusters.

```{r, eval=FALSE, warning=FALSE, error=FALSE}

finalCluster <- myCluster$cluster
finalCluster

setwd("/Users/CheHoon/Desktop/SP2022/DA350/Lab3/clusters")

for (k in 1:25) {
  dir.create(paste("Cluster", k, sep=""))  
}

for (i in 1:4410) {
  clusterNum = finalCluster[i]
  setwd(paste("/Users/CheHoon/Desktop/SP2022/DA350/Lab3/clusters/Cluster",clusterNum,sep = ""))
  writePNG(hieroglyphs[[i]], files[i])
}


```


### Question 7.

We predict which clusters new hieroglyph images will be classified. We find the mean value of each pixel for each cluster in order to find the centers. We then find the euclidean distance between the centers and the new image, and classify it to the cluster with the minimum distance.

```{r, eval=FALSE, warning=FALSE, error=FALSE}

setwd("/Users/CheHoon/Desktop/SP2022/DA350/Lab3/Predictions")
files_predict = list.files()
hieroglyphs_predict <- lapply(files_predict, function(x) readPNG(x))
predictions_df <- data.frame(do.call(rbind, lapply(hieroglyphs_predict, function(x) t(matrix(x)))))


predictCluster <- function(imagesCopy, predictions_df){
  
  predictionList <- numeric(nrow(predictions_df))
  imagesCopy$cluster <- myCluster$cluster
  imageCenter <- as.data.frame(imagesCopy %>% group_by(cluster) %>% summarise_all("mean"))
  imageCenter <- subset(imageCenter, select = -c(cluster))
  
  
  distanceList <- numeric(25)
  for (i in 1:10) {
    
    for (j in 1:25) {
      
      mySum = 0
      
      for (k in 1:3750) {
        
        mySum = mySum + (imageCenter[j,k]-predictions_df[i,k])^2
        
      }
      
      distanceList[j] = sqrt(mySum)
      
    }
    print(paste("Prediction image number : ", i, sep=""))
    minVal = min(distanceList)
    predictionCluster = which(distanceList == minVal)
    print(paste("Cluster : ", predictionCluster, sep=""))
    predictionList[i] = predictionCluster
  }
  
  return(predictionList)
  
}



```





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


